### 2021-01-11 tex/final-presentation.tex
Finish presentation.

  \begin{frame}{Outline}
    \begin{itemize}
      \item Motivation
      \item Model-Driven Architecture
      \item Low-Code Architecture
      \item Criticisms
      \item Planned Project Procedure
    \end{itemize}
  \end{frame}

@@

  \begin{frame}{Open Standard Business Platform (OSBP)}
    \begin{itemize}
      \item open-source
      \item plug-in for the Eclipse IDE developed since 2016
      \item community version of the commercial OS.bee product developed by COMPEX
      \item latest version over one year old
      \item does not work with latest version of the Eclipse IDE
    \end{itemize}
  \end{frame}

@@

@@

@@

  \begin{frame}{Mendix}
    \begin{itemize}
      \item commercial
      \item founded in 2005 as a subsidiary of Siemens
      \item web-based platform (Mendix Studio) and Windows application \\
            (Mendix Studio Pro) with advanced features
      \item test by signing up for a regular account which allows hosting unlimited applications (with 1GB of memory and 0.5GB of storage per application)
    \end{itemize}
  \end{frame}
\end{document}

>>>>

  \begin{frame}{Outline}
    \begin{itemize}
      \item Motivation
      \item Model-Driven Architecture
      \item Low-Code Architecture
      \item Criticisms
      \item Evaluation of Low-Code Tools
      \item Findings \& Future Work
      \item Conclusion
    \end{itemize}
  \end{frame}

@@

  \begin{frame}{Open Standard Business Platform (OSBP)}
    \begin{itemize}
      \item open-source
      \item plug-in for the Eclipse IDE developed since 2016
      \item community version of the commercial OS.bee product \\
            developed by COMPEX
      \item latest version over one year old
      \item does not work with latest version of the Eclipse IDE
    \end{itemize}
  \end{frame}

@@

  \begin{frame}{Corteza Low Code}
    \begin{itemize}
      \item made for building record-based management applications
      \item process for building the test application:
        \begin{enumerate}
          \item create application namespace
          \item create module for TODO List records, specifying the necessary fields (status, title, body)
          \item create page and add a list block linked to the module
        \end{enumerate}
    \end{itemize}
  \end{frame}

  \begin{frame}[standout]{Corteza Low Code: Editor}
    \includegraphics[width=\textwidth]{corteza-low-code-pages}
  \end{frame}

  \begin{frame}[standout]{Corteza Low Code: Application}
    \begin{figure}
      \includegraphics[width=\textwidth]{corteza-low-code-crm-home}
    \end{figure}
  \end{frame}

@@

  \begin{frame}{Oracle APEX (Application Express)}
    \begin{itemize}
      \item process for building the test application:
        \begin{enumerate}
          \item create database table for TODO List items \\
                (requires basic SQL knowledge)
          \item create new blank application
          \item add list view to home page and select \\
                the corresponding database table
          \item add new form page (dialog style) and select \\
                the corresponding database table
          \item add a button to the home page that opens the form page
          \item add a dynamic action that refreshes the list view \\
                when the form dialog is closed
        \end{enumerate}
    \end{itemize}
  \end{frame}

  \begin{frame}[standout]{Oracle APEX: Page Designer}
    \includegraphics[width=\textwidth]{oracle-apex-page-designer}
  \end{frame}

  \begin{frame}[standout]{Oracle APEX: Application}
    \includegraphics[width=\textwidth]{oracle-apex-todo-list-application}
  \end{frame}

@@

  \begin{frame}{Simplifier}
    \begin{itemize}
      \item process for building the test application:
        \begin{enumerate}
          \item create database connector (SQLite)
          \item create database schema for TODO List items and deploy to connector
          \item add list view and button to home page
          \item add new page containing a form with input fields and button
          \item create processes for
            \begin{itemize}
              \item loading items into list view
              \item submitting the form page
              \item opening the form page with the button
            \end{itemize}
        \end{enumerate}
    \end{itemize}
  \end{frame}

  \begin{frame}[standout]{Simplifier: Process}
    \includegraphics[width=\textwidth]{simplifier-story-overview}
  \end{frame}

  \begin{frame}[standout]{Simplifier: Process}
    \includegraphics[width=\textwidth]{simplifier-story-ui-action}
  \end{frame}

  \begin{frame}[standout]{Simplifier: Application Editor}
    \includegraphics[width=\textwidth]{simplifier-application-editor}
  \end{frame}

  \begin{frame}[standout]{Simplifier: Application}
    \includegraphics[width=\textwidth]{simplifier-shopping-list-application}
  \end{frame}

  \begin{frame}{Mendix}
    \begin{itemize}
      \item commercial
      \item founded in 2005 as a subsidiary of Siemens
      \item web-based platform (Mendix Studio) and Windows application \\
            (Mendix Studio Pro) with advanced features
      \item test by signing up for a regular account which allows hosting unlimited applications (with 1GB of memory and 0.5GB of storage per application)
    \end{itemize}
  \end{frame}

  \begin{frame}{Mendix}
    \begin{itemize}
      \item process for building the test application:
        \begin{enumerate}
          \item create blank application
          \item add list view and button to home page, which prompts the user to select or create a new data source
          \item create data source for TODO List items
          \item add new form page linked to the corresponding data source
          \item add button to home page that opens the form page
        \end{enumerate}
    \end{itemize}
  \end{frame}

  \begin{frame}[standout]{Mendix Studio}
    \includegraphics[width=\textwidth]{mendix-studio}
  \end{frame}

  \begin{frame}[standout]{Mendix: Application}
    \includegraphics[width=\textwidth]{mendix-todo-list-application}
  \end{frame}

  \begin{frame}{Unique Features}
    \begin{itemize}
      \item \textbf{Oracle APEX}
        \begin{itemize}
          \item Creation of applications from existing data, e.g. CSV files
        \end{itemize}
      \item \textbf{Mendix}
        \begin{itemize}
          \item AI-assisted wizard for creating custom workflows (microflows)
          \item publishing as native mobile applications for iOS and Android
        \end{itemize}
    \end{itemize}
  \end{frame}

  \begin{frame}{Assessment of Low-Code Tools}
    \begin{itemize}
      \item ease of use
      \item customisability
      \item portability
      \item scalability
      \item suitability for mission-critical applications
    \end{itemize}
  \end{frame}

  \begin{frame}{Assessment of Low-Code Tools}
    \begin{table}[]
    \begin{tabular}{l|l|l|l|l|l|}
    \cline{2-6}
                                           & \textbf{OSBP} & \textbf{Corteza} & \textbf{APEX}   & \textbf{Simplifier} & \textbf{Mendix} \\ \hline
    \multicolumn{1}{|l|}{ease of use}      & N/A  & high    & medium & medium     & high   \\ \hline
    \multicolumn{1}{|l|}{customisability}  & N/A  & low     & medium & medium     & high   \\ \hline
    \multicolumn{1}{|l|}{portability}      & N/A  & medium  & low    & medium     & medium \\ \hline
    \multicolumn{1}{|l|}{scalability}      & N/A  & medium\footnote{medium in general, high for certain types of applications, e.g. management applications} & high   & high       & high   \\ \hline
    \multicolumn{1}{|l|}{mission-criticality} & low  & high    & high   & medium\footnote{medium only due to the problems encountered, high otherwise}    & high   \\ \hline
    \end{tabular}
    \end{table}
  \end{frame}

  \begin{frame}{Findings \& Future Work}
    \begin{itemize}
      \item very small number of open-source low-code tools
        \begin{itemize}
          \item assumption: open-source community mostly consists of developers, so there is no need/demand for low-code platforms
          \item future work: investigate low-code other types of platforms and compare their presence in the open-source vs. the commercial space
        \end{itemize}
      \item more streamlined user interface in the more mature products like Oracle APEX and Mendix
      \item lack of portability is a valid concern for low-code tools, \\
            stored data may be the only element of a platform that is portable by using
    \end{itemize}
  \end{frame}

  \begin{frame}{Conclusion}
    \begin{itemize}
      \item low-code platforms are a valid alternative to model-driven development
      \item low-code platforms are not as flexible and limited in the ways they can be extended
      \item choice between low-code and model-driven architecture:
        \begin{itemize}
          \item highly dependent on which low-code platform is used, \\
                similar to choosing a programming language or framework for a particular task
          \item highly dependent on the application requirements, \\
                low-code well-suited for management applications (e.g. process management, CRM)
        \end{itemize}
    \end{itemize}
  \end{frame}

  \begin{frame}[standout]
    Questions?
  \end{frame}
\end{document}

### 2021-01-11 tex/paper.tex
Add some slides.
Finish presentation.

Simplifier is, like Oracle APEX, a web-based application development environment on low-code basis. Simplifier offers a demo platform called “Simplifier Playground”, an instance of the Simplifier platform specifically set up for trialling the platform. The trial instance is reset to defaults every night, so all data stored on it will not be persisted. They also offer a “Freemium” instance, available by registering via an email address, which has no such limitation, we therefore applied for such an instance. Shortly afterwards, we received a confirmation email with the URL to our instance, which unfortunately returned a “400 Bad Request” error. We contacted the Simplifier Support Team about this and decided to get familiar with the platform using the Simplifier Playground in the meantime.

@@

The Corteza-Low-Code platform is spreadsheet based, making is very easy to use. Creating a simple TODO List application was possible in less than 10 minutes and should be possible in less than 5 minutes once you are familiar with the platform. Apart from choosing the order of predefined building blocks, customising the layout of an application is hindered by the fact that it is spreadsheet based. The selection of pre-built blocks is very limited, as is the selection of colour schemes. Corteza uses a MySQL database in the back end, making at least the stored data very portable. The created applications are virtually unportable since they are tightly integrated into the Corteza platform, custom scripts are at least partially portable since they are written in JavaScript. In terms of scalability, creating complex applications will be difficult using the small selection of building blocks. Process management applications which don't rely on many different UI elements, like the included Corteza CRM, can definitely be implemented in a scalable manner. Given the overall simplicity of the Corteza platform, it is certainly suitable for creating mission-critical applications.

>>>>

Simplifier is, like Oracle APEX, a web-based application development environment on low-code basis. Simplifier offers a demo platform called “Simplifier Playground”, an instance of the Simplifier platform specifically set up for trialling the platform. The trial instance is reset to defaults every night, so all data stored on it will not be persisted. They also offer a “Freemium” instance, available by registering via an email address, which has no such limitation, we therefore applied for such an instance. Shortly afterwards, we received a confirmation email with the URL to our instance, which unfortunately returned a “400 Bad Request” error. We contacted the Simplifier support team about this and decided to get familiar with the platform using the Simplifier Playground in the meantime.

@@

The Corteza-Low-Code platform is meant for building record-based management applications, making is very easy to use. Creating a simple TODO List application was possible in less than 10 minutes and should be possible in less than 5 minutes once you are familiar with the platform. Apart from choosing the order of predefined building blocks, customising the layout of an application is hindered by the fact that it is specialised for building management applications. The selection of pre-built blocks is very limited, as is the selection of colour schemes. Corteza uses a MySQL database in the back end, making at least the stored data very portable. The created applications are virtually unportable since they are tightly integrated into the Corteza platform, custom scripts are at least partially portable since they are written in JavaScript. In terms of scalability, creating complex applications will be difficult using the small selection of building blocks. Process management applications which don't rely on many different UI elements, like the included Corteza CRM, can definitely be implemented in a scalable manner. Given the overall simplicity of the Corteza platform, it is certainly suitable for creating mission-critical applications.
